The SMT Competition 2015--2018

### 2018-11-21 2018/report/SMT-COMP.tex
Conclusion

StarExec typically reserves between~80 and~150 machines for the SMT
Competition, allowing its workload to be completed within two to three
weeks.  By these measures, the SMT Competition has become not only one
of the heaviest users of StarExec, but also one of the largest
competitions in automated reasoning.

@@

A contentious issue with this ranking has been the scoring of
incorrect solver responses.  In the division rankings, an incorrect
response causes a solver to be ranked behind all correct solvers for
that division.  But the competition's main track consists of many
independent divisions.  To balance the strong interest in correct
solvers against the risk of stifling innovation, the competition-wide
ranking was revised in~2017 to use a fixed numeric penalty for each
division with incorrect responses.  The value of this penalty is
currently set so that entering a (possibly buggy) solver that can
solve all benchmarks into a division has a positive expected score if
the probability of a soundness bug in the solver is
below~$20\,\%$.~\cite{rules18}

@@

inclusion of check-sat commands with unknown status in the application track

competition-wide rankings for application and unsat-core track

more benchmarks, especially for underrepresented (i.e., small) logics

The StarExec infrastructure has been vitally important in allowing the
competition to grow to its current size.  The competition organizers
have given the StarExec developers comprehensive feedback, and filed
dozens of feature requests and bug reports since~2013.  The
infrastructure has become increasingly stable; serious issues are now
rare.  When they did surface during ongoing competitions, they
received a swift response by the StarExec team (under the lead of
Aaron Stump).  This has been critical to the success of the
competition on more than one occasion.

>>>>

StarExec typically reserves between~80 to~150 machines for the SMT
Competition, allowing its workload to be completed within two to three
weeks.  By these measures, the SMT Competition has become not only one
of the heaviest users of StarExec, but also one of the largest
competitions in automated reasoning.

@@

A contentious issue with this ranking has been the scoring of
incorrect solver responses.  In the division rankings, an incorrect
response causes a solver to be ranked behind all correct solvers for
that division.  But the competition's main track consists of many
independent divisions.  To balance the strong interest in correct
solvers against the risk of stifling innovation, the competition-wide
ranking was revised in~2017 to use a fixed numeric penalty for each
division with incorrect responses.  The value of this penalty is
currently set so that entering a (possibly buggy) solver that can
solve all benchmarks into a division has a positive expected score if
the probability that a soundness bug in the solver will be triggered
is below~$20\,\%$.~\cite{rules18}

@@

After 13 years, the SMT Competition is a well-established event that
garners wide interest from developers and users of SMT solvers.  The
competition continues to contribute to its original goals, including
adoption of the SMT-LIB format and sparking further advances in SMT,
and it provides a yardstick of performance for SMT solvers.

The last FLoC Olympiad was a period of significant growth for the SMT
Competition.  The average number of solver submissions increased
from~12 (for 2005--2014) to~26 (for 2015--2018).  The number of
SMT-LIB benchmarks used in the competition increased from~77,352
in~2014~\cite{CDW14} to 342,498 in 2018.  This is partly because new
benchmarks were added to the SMT Library, but also because benchmarks
with unknown status are now eligible for the competition, and because
the earlier culling of benchmarks was abandoned.  Since the
re-introduction of the unsat-core track in~2016, the competition again
has three separate tracks.  This, together with new logics in the SMT
Library, has brought the total number of competition divisions to 115
in~2018, up from just 40 divisions in~2014~\cite{CDW14}.  The number
of job pairs increased from 339,714 in~2014 to~1,776,062~in 2018, and
the total wall-clock time in~2018 exceeded 7.6~years.  By these
measures, the SMT Competition is one of the largest competitions in
automated reasoning.

The StarExec infrastructure has been vitally important in allowing the
competition to grow to its current size.  The competition organizers
have given the StarExec developers comprehensive feedback, and filed
dozens of feature requests and bug reports since 2013.  The
infrastructure has become increasingly stable; serious issues are now
rare.  When they did surface during ongoing competitions, they
received a swift response by the StarExec team (under the lead of
Aaron Stump).  This has been critical to the success of the
competition on more than one occasion.

Despite the substantial amount of computational resources available to
the competition through StarExec, the competition has reached a size
where future organizers will have to carefully consider the number of
participants and benchmarks each year, to determine whether the
current time limits remain tenable.  (Otherwise, reducing the time
limits would probably be a better solution than reverting to the
pre-2014 practice of benchmark culling.)

Further development of the competition should be undertaken with care,
keeping in mind that any major changes require not just the
competition organizers, but also participants to adapt.  Immediate
action items for next year's organizers are
\begin{itemize}
\item the inclusion of {\tt check-sat} commands with unknown status in
  the application track; and
\item the definition of competition-wide rankings for the application
  and unsat-core track.
\end{itemize}

For the inclusion of new tracks or divisions, a two-phase approach has
proven successful.  A new division is first declared
\emph{experimental}, meaning that results will be reported but no
official winner will be announced.  This allows competition organizers
and participants to gain valuable experience, while reducing the
impact of mistakes.  If successful, the division may then become
regular in the following year.

Organizing the competition is a non-trivial amount of work.  Much of
this happens behind the scenes.  For instance, the competition tools
(such as the benchmark scrambler and trace executor) often require
updating when new divisions are added, or when changes are made to
StarExec.  One could imagine adding further tracks (e.g., for
proof-producing SMT solvers, or for variants such as max-SMT) to the
competition, but the effort required to support such tracks will need
to be balanced against their added value.

To ensure that the competition results are practically relevant, it is
important that the SMT Library contains benchmarks that cover a wide
range of applications.  To this end, we again encourage users of SMT
solvers to submit new and interesting benchmarks to the library,
especially for underrepresented logics.  This is not just a valuable
contribution to the community, but also---since these benchmarks will
then be used in the competition---a straightforward way to get solver
developers to tune their solvers accordingly.

\paragraph{Acknowledgments}

We would like to thank Aaron Stump for allowing the SMT Competition to
run on StarExec; the SMT-LIB maintainers, namely Clark Barrett, Pascal
Fontaine, Aina Niemetz and Mathias Preiner, for preparing the annual
SMT-LIB release; all contributors of new benchmarks to the SMT
Library; and last but not least all solver developers who entered
their solver into the competition.
